Deep Learning

### 2015-03-24 deepreview.tex
modified figures and captions (1 and 2) following suggestions of referee 1

% deep learning: learning representations
Deep learning is a set of methods that allows a machine to be fed with
raw data and to automatically discover suitable representations of the
data. The name comes from the fact that deep learning machines are
multilayer stacks of trainable modules, each of which turns the
representation of the data in its input into a more abstract and
higher-level representation on its output in which relevant
characteristics and amplified and irrelevant variabilities are
attenuated. For example, an image comes in the form of an array of
pixel values, features in the first layer would represent the presence
or absence of edges of a particular orientation at all locations in
the image, the second layer would  detect motifs by spotting particular
arrangements of edges, regardless of small variations in the positions
of the edges.  The third layer would assemble motifs into parts of
object, and subsequent layers would to detect objects as combinations of
parts. The key aspect of deep learning is that {\em these features and
  representations are not designed by human engineers, but are
  automatically learned from data.}

@@

\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\fbox{\includegraphics[width=0.69\linewidth]{netvis-simple-2S.png}}
\fbox{\includegraphics[width=0.28\linewidth]{chain-rule.pdf}}}
\centerline{
\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
}
\caption{
{\bf Top left.} A multi-layer neural network can distort the
input space to make the classes (whose examples are supposed to
be on the red and blue lines, here) linearly separable. Note how
a regular grid in input space is also transformed. This is an
illustrative example with only two input units, two hidden units
and one output unit, but the networks used for object recognition
or natural language processing contain tens or hundreds of thousand
of units. Reproduced with permission by Chris Olah from~\url{http://colah.github.io/}.
\newline
{\bf Top right.} The chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). 
\newline
{\bf Bottom left.} 
The equations used for computing the forward pass in a neural
net with two hidden layers. At each layer, we first compute the total
input $z$ to each unit, which is a weighted sum of the outputs of the
units in the layer below. Then a non-linear function $f(.)$ is applied
to $z$ to get the output of the unit.  For simplicity, we have omitted
bias terms. The non-linear function used in neural networks include 
the rectified linear unit (ReLU) $f(z) = max(0, z)$, commonly-used in 
recent years, as well as the more traditional sigmoids, such as the 
hyberbolic tangent, $f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ 
and logistic function logistic, $f(z) = 1/(1+\exp(-z))$.
{\bf Bottom right.} 
The equations used for computing the backward pass. At each
hidden layer we compute the error-derivative w.r.t. the output of each
unit which is a weighted sum of the error-derivatives w.r.t. the total
inputs to the units in the layer above. We then convert the
error-derivative w.r.t. the output into the error-derivative w.r.t.
the input by multipling it by the gradient of $f(z)$.  At the output
layer, the error-derivative w.r.t. the output of a unit is computed by
differentiating the cost function. This gives $y_l-t_l$ if the cost
function for unit $l$ is $0.5(y_l-t_l)^2$ where $t_l$ is the target
value. Once the $\partial E/\partial z_k$ is known, the
error-derivative for the weight $w_{jk}$ on the connection from unit
$j$ in the layer below is just $y_j \partial E/\partial z_k$.
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

\begin{figure}
\begin{center}
  \vspace{0.2cm}
  \includegraphics[width=0.70\linewidth]{convnet-diagram}
\end{center}
\vspace{-0.6cm}
\caption{The outputs of each layer (column) and feature map (rectangular image)
of a typical convolutional network architecture
applied on the image of a handwritten digit (on the top left). Each rectangular
image is a feature map, corresponding to one of the learned features 
detected at each of the image positions.
Each column corresponds to a layer, with information flowing from the left to
the right. We see that lower level features mostly act as oriented edge detectors.
}
%\vspace{-0.8cm}
\label{fig:convnet}
\end{figure}

>>>>

% deep learning: learning representations
Deep learning is a set of methods that allows a machine to be fed with
raw data and to automatically discover suitable representations of the
data. The name comes from the fact that deep learning machines are
multilayer stacks of trainable modules, each of which (e.g. a neural
network layer) turns the
representation of the data in its input into a more abstract and
higher-level representation on its output in which relevant
characteristics and amplified and irrelevant variabilities are
attenuated. For example, an image comes in the form of an array of
pixel values, features in the first layer would represent the presence
or absence of edges of a particular orientation at all locations in
the image, the second layer would  detect motifs by spotting particular
arrangements of edges, regardless of small variations in the positions
of the edges.  The third layer would assemble motifs into parts of
object, and subsequent layers would to detect objects as combinations of
parts. The key aspect of deep learning is that {\em these features and
  representations are not designed by human engineers, but are
  automatically learned from data.}

@@

\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\fbox{\includegraphics[width=0.69\linewidth]{netvis-simple-2S.png}}
\fbox{\includegraphics[width=0.28\linewidth]{chain-rule.pdf}}}
\centerline{
\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
}
\caption{
{\bf Top left.} A multi-layer neural network can distort the
input space to make the classes (whose examples are supposed to
be on the red and blue lines, here) linearly separable. Note how
a regular grid in input space is also transformed. This is an
illustrative example with only two input units, two hidden units
and one output unit, but the networks used for object recognition
or natural language processing contain tens or hundreds of thousand
of units. Reproduced with permission by Chris Olah from~\url{http://colah.github.io/}.
\newline
{\bf Top right.} The chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). 
\newline
{\bf Bottom left.} 
The equations used for computing the forward pass in a neural
net with two hidden layers and one output layer, each constituting
a module through which one can backpropagate gradients. At each layer, we first compute the total
input $z$ to each unit, which is a weighted sum of the outputs of the
units in the layer below. Then a non-linear function $f(.)$ is applied
to $z$ to get the output of the unit.  For simplicity, we have omitted
bias terms. The non-linear function used in neural networks include 
the rectified linear unit (ReLU) $f(z) = max(0, z)$, commonly-used in 
recent years, as well as the more traditional sigmoids, such as the 
hyberbolic tangent, $f(z) = (\exp(z)-\exp(-z))/(\exp(z)+\exp(-z))$ 
and logistic function logistic, $f(z) = 1/(1+\exp(-z))$.
{\bf Bottom right.} 
The equations used for computing the backward pass. At each
hidden layer we compute the error-derivative w.r.t. the output of each
unit which is a weighted sum of the error-derivatives w.r.t. the total
inputs to the units in the layer above. We then convert the
error-derivative w.r.t. the output into the error-derivative w.r.t.
the input by multipling it by the gradient of $f(z)$.  At the output
layer, the error-derivative w.r.t. the output of a unit is computed by
differentiating the cost function. This gives $y_l-t_l$ if the cost
function for unit $l$ is $0.5(y_l-t_l)^2$ where $t_l$ is the target
value. Once the $\partial E/\partial z_k$ is known, the
error-derivative for the weight $w_{jk}$ on the connection from unit
$j$ in the layer below is just $y_j \partial E/\partial z_k$.
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

\begin{figure}
\begin{center}
  \vspace{0.2cm}
  \includegraphics[width=0.70\linewidth]{convnet-diagram-labeled}
\end{center}
\vspace{-0.6cm}
\caption{The {\em outputs} (not the filters) of each layer (each column of the figure) 
of a typical convolutional network architecture
applied on the image of a handwritten digit (left). Each rectangular
image is a feature map, corresponding to the output for one of the learned features 
detected at each of the image positions.
Each column corresponds to a layer, with information flowing from the left to
the right. We see that lower level features mostly act as oriented edge detectors.
}
%\vspace{-0.8cm}
\label{fig:convnet}
\end{figure}
